## MDD DNAm score in ALSPAC

### Background

A DNAm model of MDD has been generated in 1824 controls and 1223 cases
> Barbu, M.C., Shen, X., Walker, R.M. et al. Epigenetic prediction of major depressive disorder. Mol Psychiatry 26, 5112–5123 (2021). 

In independent data it has explained 1.75% of variance in MDD.

### Results in ALSPAC

We used ALSPAC DNAm profiles for the study children from ages 15 and 24 years
and from their mothers during pregnancy (mean ~29 years old) and approximately 18 years later.
Depression was assessed around the same time as the DNAm profiles.

Around age 15, depression was determined using
DAWBA (whether derpessive disorder present by self-report computer prediction, ICD-10 and DSM-IV: fh6892)
and CIS-R (any ICD-10 diagnosis of depression, mild, moderate or severe: fjci1001). 

Around age 24, depression was determined using the
CIS-R (mild depression or not: fkdq1000;
moderate depression or not: fkdq1010;
severe depression numbers too low).

Depression in mothers was determined using the EPDS (>= 13)
during pregnancy (at 32 weeks gestation: c600) and
approximately 18 years later (t3255).

```{r}
depression.vars <- list(aries.15=c("dawba.15","cisr.17"),
                        aries.24=c("cisr.mild.24","cisr.mod.24"),
                        aries.preg="epds.preg",
                        aries.mid="epds.mid")

stats <- data.frame(dataset=rep(names(depression.vars), sapply(depression.vars,length)),
                    variable=unlist(depression.vars),
                    stringsAsFactors=F)
```

```{r}
library(pROC)
for (i in 1:nrow(stats)) {
    var <- stats$variable[i]
    ds <- stats$dataset[i]
    if (all(get(ds)$samples$alnqlet %in% alspac.table$alnqlet))
        idx <- match(get(ds)$samples$alnqlet, alspac.table$alnqlet)
    else
        idx <- match(get(ds)$samples$aln, alspac.table$aln)
    depression <- alspac.table[[var]][idx]
    if (is.logical(depression)) depression <- sign(depression)
    stats$r[i] <- cor(depression, scores[[ds]], use="p")
    stats$r2[i] <- stats$r[i]^2
    stats$p[i] <- cor.test(depression, scores[[ds]])$p.value
    stats$auc[i] <- auc(depression, scores[[ds]])
    stats$n[i] <- length(which(!is.na(depression) & !is.na(scores[[ds]])))
    stats$n.cases[i] <- sum(depression,na.rm=T)
    age <- alspac.table[[paste(var, "age", sep=".")]][idx]
    na.idx <- which(is.na(depression) | is.na(scores[[ds]]))
    if (length(na.idx) > 0)
        age[na.idx] <- NA
    stats$age.mean[i] <- mean(age,na.rm=T)
    stats$age.sd[i] <- sd(age,na.rm=T)
    stats$age.min[i] <- min(age, na.rm=T)
    stats$age.max[i] <- max(age, na.rm=T)
    plots.boxplot(score ~ depression, data.frame(score=scores[[ds]], depression=depression),
                  main=paste(var, " vs mdd(", ds, ")", sep=""),
                  xlab=var,
                  ylab="MDD DNAm score")
}
rownames(stats) <- NULL

kable(stats,digits=4)
```
